---
title: "Simulated PWMs + IC values"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
---


__Author:__ Amanda Everitt 


```{r, warning=FALSE}
suppressPackageStartupMessages({
  library(TFBSTools)
  library(dplyr)
  library(ggplot2)
  library(tidyr)
  library(ComplexHeatmap)
  library(circlize)
  library(gridExtra)
  library(knitr)
  library(RColorBrewer)
  library(monaLisa)
  library(BSgenome.Hsapiens.UCSC.hg38)
  library(seqLogo)
})

setwd("../09_simPWMs/00_constructing_PWMs/") 

#https://bioconductor.org/packages/release/bioc/vignettes/TFBSTools/inst/doc/TFBSTools.html#dynamic-random-profile-generation
```

# JASPAR PWMs
## 1. Read in jaspar db and check the matrices look we expect. 
### (showing a random motif here)

```{r}
jaspar_pcm <- TFBSTools::readJASPARMatrix("../../05_footprinting/program_input_files/JASPAR2022_CORE_vertebrates_non-redundant_pfms_jaspar.txt", matrixClass=c("PFM"))
names(jaspar_pcm) = sapply(jaspar_pcm, function(x) paste0(toupper(x@name),"_", x@ID))

jaspar_ppm <- TFBSTools::toPWM(jaspar_pcm, pseudocounts=0.8, type = "prob")

jaspar_pwm <- TFBSTools::toPWM(jaspar_pcm, pseudocounts=0.8, type = "log2probratio")

jaspar_icm <- toICM(jaspar_pcm, pseudocounts=0.8, schneider=TRUE)

seqLogo(jaspar_ppm$ARNT2_MA1464.1@profileMatrix)
print("pcm"); jaspar_pcm$ARNT2_MA1464.1@profileMatrix
print("ppm"); jaspar_ppm$ARNT2_MA1464.1@profileMatrix
print("pwm"); jaspar_pwm$ARNT2_MA1464.1@profileMatrix
print("icm"); jaspar_icm$ARNT2_MA1464.1@profileMatrix
```



```{}
#TEST CODE FOR SIMULATIONS. 
#train dirichlet multinomial mixture model with max 6 components, 
#used trained parameters to sample from matrices and create PPMs (though called PWMs here)

dmmParameters <- dmmEM(jaspar_pcm[1:5], K=6, alg="C")
ppmSampled <- rPWMDmm(jaspar_pcm[1:5], dmmParameters$alpha0, dmmParameters$pmix,
                      N=1, W=5)
ppmSampled[[1]]
```

### First road block:
- The first thing I noticed is that the simulated results are returned as PPMs, but the ICM function only works on PCM matrices. Initially this seemed confusing, but its because the ICM function needs to know the total number of observations to do the schneider correction. 
- Instead, lets calculate the IC ourselves from the ppms and add a constant observation value for the simulated ones. 
- Code block below checks that my hand-calculated IC values and the programs are identical. 

```{r}
## THAT is why their ICM function wants the PCMs -- its for the schnieder correction that's based on the total observations. 

mylist = list()

for (tf in names(jaspar_icm)){
  icm_mat = jaspar_icm[[tf]]@profileMatrix
  pcm_mat = jaspar_pcm[[tf]]@profileMatrix
  ppm_mat = jaspar_ppm[[tf]]@profileMatrix
  
  n_observed = colSums(pcm_mat)[1]
  n_len = ncol(icm_mat)
    
  h = sum(ppm_mat * log2((ppm_mat/0.25)))
  if (n_observed <= 30){
    e = (2-TFBSTools:::schneider_Hnb_precomputed(n_observed)) * n_len
  }else{
    e = (2-TFBSTools:::schneider_Hnb_approx(n_observed, 2)) * n_len
  }
  
  mylist[[tf]] = c(sum(icm_mat), #ic 
                   n_len, #len
                   n_observed, #obs
                   h-e, #ic by hand. for no correction remove e.
                   mean(ppm_mat["G",] + ppm_mat["C",]),
                   max(mean(ppm_mat["A",]), mean(ppm_mat["G",]), 
                       mean(ppm_mat["C",]), mean(ppm_mat["T",])))
}


ic_df = as.data.frame(do.call("rbind",mylist))
colnames(ic_df) = c("IC","len","obs","IC_byHand","GC_content", "Repeat_content")
ic_df$ic_len_norm = ic_df$IC/ic_df$len #check these are below 2. 
ic_df$TF = rownames(ic_df)
head(ic_df)
```

```{r}
#some examples of the extremes
seqLogo(jaspar_ppm$SP1_MA0079.5@profileMatrix)
seqLogo(jaspar_ppm$ATF3_MA0605.2@profileMatrix)
seqLogo(jaspar_ppm$ZNF384_MA1125.1@profileMatrix)
```

```{r, fig.height=6, fig.width=15}
grid.arrange(
ggplot(ic_df, aes(x=len)) + 
    geom_histogram(color="black",fill="cornflowerblue", bins = 30) + theme_bw() +
  ggtitle("Length of 841 motifs"), 

ggplot(ic_df, aes(x=IC)) + 
    geom_histogram(color="black",fill="cornflowerblue", bins = 30) + theme_bw() +
  ggtitle("IC value"), 

ggplot(ic_df, aes(x=ic_len_norm)) + 
    geom_histogram( color="black",fill="cornflowerblue", bins = 30) + theme_bw() +
  ggtitle("IC value divided by length"), 

ggplot(ic_df, aes(x=obs)) + 
    geom_histogram( color="black",fill="cornflowerblue", bins = 30) + theme_bw() +
  ggtitle("Number of observations"), 

ggplot(ic_df, aes(x=GC_content)) + 
    geom_histogram( color="black",fill="cornflowerblue", bins = 30) + theme_bw() +
  ggtitle("GC content"), 

ggplot(ic_df, aes(x=Repeat_content)) + 
    geom_histogram( color="black",fill="cornflowerblue", bins = 30) + theme_bw() +
  ggtitle("Repeat content"),

ncol=3
)
```

```{r, fig.width=10, fig.height=10}
grid.arrange(
  ggplot(ic_df, aes(x=len, y=IC, color="FALSE")) + 
    geom_point() + theme_bw() +
    scale_color_manual(values = "black", name="fake") + 
    ggtitle("PWM length by IC value"),
  
  ggplot(ic_df, aes(x=GC_content, y=IC, color="FALSE")) + 
    geom_point() + theme_bw() +
    scale_color_manual(values = "black", name="fake") + 
    ggtitle("PWM GC content by IC value"),
  
  ggplot(ic_df, aes(x=GC_content, y=Repeat_content, color="FALSE")) + 
    geom_point() + theme_bw() +
    scale_color_manual(values = "black", name="fake") + 
    ggtitle("GC content by Repeat Content"),
  
  ncol=2
)

```


### Calculate genome prevalance of chr10. 
```{r, eval=FALSE, warning=FALSE}
seqs <- BSgenome::getSeq(BSgenome.Hsapiens.UCSC.hg38, "chr10")

res <- findMotifHits(query = jaspar_pwm,
                     subject = seqs,
                     min.score = 6.0, 
                     method = "matchPWM",
                     BPPARAM = BiocParallel::MulticoreParam(workers=10)
                     #BPPARAM = BiocParallel::SerialParam()
                     )

res$fullname = paste0(toupper(res$pwmname), "_", res$pwmid)
tmp_df = data.frame(table(res$fullname))
colnames(tmp_df) = c("TF", "chr10_prevalence")

merged_df <- merge(ic_df, tmp_df, by="TF", all=TRUE) 

write.csv(merged_df[, c("TF", "IC", "len", "obs", "GC_content", "Repeat_content", "chr10_prevalence")], file="jaspar2022_ICcontent.csv", quote=F, row.names = F)
```

```{r}
merged_df = read.csv("jaspar2022_ICcontent.csv")
head(merged_df)
```

```{r, fig.width=12, fig.height=12}
model1 = lm(merged_df$chr10_prevalence ~ poly(merged_df$len, 2, raw=TRUE))
model2 = lm(merged_df$chr10_prevalence ~ poly(merged_df$IC, 2, raw=TRUE))
model3 = lm(merged_df$chr10_prevalence ~ poly(merged_df$GC_content, 2, raw=TRUE))
model4 = lm(merged_df$chr10_prevalence ~ poly(merged_df$Repeat_content, 2, raw=TRUE))

grid.arrange(
  ggplot(merged_df, aes(x=chr10_prevalence, y = len, color="FALSE")) + 
    geom_point() + theme_bw() + 
    scale_color_manual(values = "black", name="fake") + 
    ggtitle(paste0("PWM length R-squared: ", round(summary(model1)$r.squared, 3))),
  
  ggplot(merged_df, aes(x=chr10_prevalence, y = IC, color="FALSE")) + 
    geom_point() + theme_bw() +
    scale_color_manual(values = "black", name="fake") + 
    ggtitle(paste0("PWM IC R-squared: ", round(summary(model2)$r.squared, 3))),
  
  ggplot(merged_df, aes(x=chr10_prevalence, y = GC_content, color="FALSE")) + 
    geom_point() + theme_bw() +
    scale_color_manual(values = "black", name="fake") + 
    ggtitle(paste0("PWM GC content R-squared: ", round(summary(model3)$r.squared, 3))),
  
  ggplot(merged_df, aes(x=chr10_prevalence, y = Repeat_content, color="FALSE")) + 
    geom_point() + theme_bw() +
    scale_color_manual(values = "black", name="fake") + 
    ggtitle(paste0("PWM Repeat content R-squared: ", round(summary(model4)$r.squared, 3))),
  
  ncol=2
)
```

```{r}
#max
seqLogo(jaspar_ppm$ZNF354A_MA1978.1@profileMatrix)
```


## Simulations
```{r}
myseed = 10
```

### 1. Simulate values from all of the jaspar db
```{r, include=FALSE}
#train dirichlet multinomial mixture model with max 6 components, 
#used trained parameters to sample from matrices and create PPMs (though called PWMs here)
dmmParameters <- dmmEM(jaspar_pcm, K=6, alg="C")
```
```{r}
#want ~59
total_n = 6
set.seed(myseed)
ppmSampled1 <- c()
for (i in seq(6, 24, 1)){
  ppmSampled1 <- c(ppmSampled1, 
                 rPWMDmm(jaspar_pcm, dmmParameters$alpha0, dmmParameters$pmix,
                      N=total_n, W=i))
}

names(ppmSampled1) = paste0("FAKE_ALL_", 1:length(ppmSampled1))
head(names(ppmSampled1))
ppmSampled1[[1]]
```

### 2. Simulate values from all jaspar clusters with > 30 members. 
```{r}
### Read in outside info
pwm_clusters = read.csv("../../05_footprinting/program_input_files/jaspar2022_motifs_to_clusters.txt", sep=" ", header = FALSE)
colnames(pwm_clusters) = c("motif_id", "cluster")
pwm_clusters$motif_id = toupper(pwm_clusters$motif_id)
rownames(pwm_clusters) = pwm_clusters$motif_id
head(pwm_clusters)
```

```{r}
dnb_domains = read.table("../../05_footprinting/program_input_files/jaspar2022_dndbd.txt", sep="[", header = FALSE)
dnb_domains$V1 =  gsub('.{1}$', '', dnb_domains$V1)
dnb_domains$V2 =  gsub('.{2}$', '', dnb_domains$V2)
dnb_domains$V3 =  gsub('.{1}$', '', dnb_domains$V3)
dnb_domains = dnb_domains %>% 
  tidyr::separate_wider_delim(V1, ",", names = c("motif", "tf"))  %>% 
  tidyr::separate_wider_delim(V2, ",", names = c("class","class2"), too_few = "align_start")  %>% 
  tidyr::separate_wider_delim(V3, ",", names = c("family","family2"), too_few = "align_start")  %>% 
  as.data.frame()
rownames(dnb_domains) = toupper(paste0(dnb_domains$tf, "_", dnb_domains$motif))

head(dnb_domains)
```

```{r, include=FALSE}
large_clust = names(table(pwm_clusters$cluster)[table(pwm_clusters$cluster) > 30])

ppmSampled2 <- c()
for (n in large_clust){
  clust = pwm_clusters[pwm_clusters$cluster == n, "motif_id"]
  sub_jaspar_pcm = jaspar_pcm[names(jaspar_pcm) %in% clust]
  dmmParameters <- dmmEM(sub_jaspar_pcm, K=6, alg="C")

  total_n = 3
  set.seed(myseed)
  internal <- c()
  for (i in seq(6, 24, 2)){
    internal <- c(internal, 
                 rPWMDmm(sub_jaspar_pcm, dmmParameters$alpha0, dmmParameters$pmix,
                      N=total_n, W=i))
  }
  names(internal) = paste0("FAKE_C", gsub("cluster", "", n), "_", 1:length(internal))
  ppmSampled2 = c(ppmSampled2, internal)
}

head(names(ppmSampled2))
```

### 3. Simulate values from USF factors. 
```{r}
usf <- read.csv("../../05_footprinting/program_input_files/USF_list.csv", header=FALSE)
```
```{r, include=FALSE}
usf_motifs = names(jaspar_pcm)[sub("\\_.*", "", names(jaspar_pcm)) %in% usf$V1]
sub_jaspar_pcm = jaspar_pcm[names(jaspar_pcm) %in% usf_motifs]
dmmParameters <- dmmEM(sub_jaspar_pcm, K=6, alg="C")
```
```{r, include=FALSE}
total_n = 3
set.seed(myseed)
ppmSampled3 <- c()
for (i in seq(6, 24, 2)){
  ppmSampled3 <- c(ppmSampled3, 
                 rPWMDmm(sub_jaspar_pcm, dmmParameters$alpha0, dmmParameters$pmix,
                      N=total_n, W=i))
}
names(ppmSampled3) = paste0("FAKE_USF_", 1:length(ppmSampled3))
head(names(ppmSampled3))
```

### 4. Join them all together and plot
```{r}
#join them all together
ppmSampled = c(ppmSampled1, ppmSampled2, ppmSampled3)
for (i in names(ppmSampled)){ rownames(ppmSampled[[i]]) = c("A", "C", "G", "T") }
```

```{r}
mylist = list()
obs = 4000 #around the median(ic_df$obs)

for (tf in names(ppmSampled)){
  ppm = ppmSampled[[tf]]
  
  h = sum(ppm * log2((ppm/0.25)))
  curr_len = ncol(ppm)
  e = (2-TFBSTools:::schneider_Hnb_approx(100, 2)) * curr_len
  
  mylist[[tf]] = c(tf,
                   h-e, #ic
                   curr_len,
                   mean(ppm["G",] + ppm["C",]),
                   max(mean(ppm["A",]), mean(ppm["G",]), 
                       mean(ppm["C",]), mean(ppm["T",]))
                   )
}

ic_df_sim = as.data.frame(do.call("rbind",mylist))
colnames(ic_df_sim) = c("TF", "IC", "len", "GC_content", "Repeat_content")
head(ic_df_sim)
```

```{r}
pcmSampled = TFBSTools:::PFMatrixList()
for (i in names(ppmSampled)){
  a <- PFMatrix(ID=i, name=i, 
                matrixClass="fake",
                bg=c(A=0.25, C=0.25, G=0.25, T=0.25),
                profileMatrix=round(ppmSampled[[i]]*4000)
                )
  #pcmSampled = c(pcmSampled, list(i = a))
  pcmSampled[[i]] = a
}
pwmSampled <- TFBSTools::toPWM(pcmSampled, pseudocounts=0.8)
```

```{r, warning=FALSE}
seqs <- BSgenome::getSeq(BSgenome.Hsapiens.UCSC.hg38, "chr10")

res <- findMotifHits(query = pwmSampled,
                     subject = seqs,
                     min.score = 6.0, 
                     method = "matchPWM",
                     BPPARAM = BiocParallel::MulticoreParam(workers=10)
                     #BPPARAM = BiocParallel::SerialParam()
                     )

res$pwmid <- as.factor(res$pwmid)
tmp_df = data.frame(tabulate(res$pwmid))
colnames(tmp_df) = c("chr10_prevalence")
tmp_df$TF = levels(res$pwmid)

merged_df <- merge(ic_df_sim, tmp_df, by="TF", all=TRUE) 
write.csv(merged_df, file="simulated_ICcontent.csv", quote=F, row.names = F)
```


```{r}
merged_df = read.csv("jaspar2022_ICcontent.csv")
merged_df$fake = FALSE
merged_df$fake_group = "not_simulated"

merged_df_sim = read.csv("simulated_ICcontent.csv")
merged_df_sim$fake = TRUE
merged_df_sim$fake_group = "simulated"
merged_df_sim$fake_group = sub("\\_.*", "", gsub("FAKE_", "", merged_df_sim$TF))

cols = c("TF", "IC", "len", "GC_content", "Repeat_content", "chr10_prevalence", "fake", "fake_group")
plot_df = rbind(merged_df[,cols], 
                merged_df_sim[,cols])

plot_df$fake_group <- as.factor(plot_df$fake_group)
plot_df$fake_group = relevel(plot_df$fake_group, ref = "not_simulated")
plot_df[c("IC", "len", "GC_content", "Repeat_content")] <- sapply(plot_df[c("IC", "len", "GC_content", "Repeat_content")],as.numeric)
```


```{r, fig.width=10, fig.height=10}
grid.arrange(
  ggplot(plot_df, aes(x=len, y=IC, color=fake)) + 
    geom_point() + 
    theme_bw() +
    scale_color_manual(values=c("black","red")),
  
  ggplot(plot_df, aes(x=len, y=IC, color=fake_group)) + 
    geom_point() + 
    theme_bw() +
    scale_color_manual(values=c("black","blue",rep("orange", 7),"red")),
  
  ncol=2
)
```

```{r, fig.width=10, fig.height=10}
grid.arrange(
  ggplot(plot_df, aes(x=GC_content, y=Repeat_content, color=fake)) + 
    geom_point() + 
    theme_bw() +
    scale_color_manual(values=c("black","red")),
  
  ggplot(plot_df, aes(x=GC_content, y=Repeat_content, color=fake_group)) + 
    geom_point() + 
    theme_bw() +
    scale_color_manual(values=c("black","blue",rep("orange", 7),"red")),
  
  ncol=2
)
```

```{r, fig.width=12, fig.height=12}
#add in the fake ones when done. 
grid.arrange(
  ggplot(plot_df, aes(x=chr10_prevalence, y = len, color=fake)) + 
    geom_point() + theme_bw() +
    scale_color_manual(values=c("black","red")),
  
  ggplot(plot_df, aes(x=chr10_prevalence, y = IC, color=fake)) + 
    geom_point() + theme_bw() +
    scale_color_manual(values=c("black","red")),
  
  ggplot(plot_df, aes(x=chr10_prevalence, y = GC_content, color=fake)) + 
    geom_point() + theme_bw() +
    scale_color_manual(values=c("black","red")),
  
  ggplot(plot_df, aes(x=chr10_prevalence, y = Repeat_content, color=fake)) + 
    geom_point() + theme_bw() +
    scale_color_manual(values=c("black","red")),
  
  ncol=2
)
```

-  remove the low length and low ic scores. 
```{r}
rem = plot_df[plot_df$fake ==TRUE & (plot_df$IC < 5 | plot_df$IC > 25), "TF"]

rem2 = plot_df[plot_df$fake ==TRUE & (plot_df$Repeat_content > 0.7), "TF"]

rem3 = plot_df[plot_df$fake ==TRUE & (plot_df$chr10_prevalence > 2.5e6), "TF"]

extremeICs = plot_df[plot_df$fake ==TRUE & (plot_df$IC < 10 | plot_df$IC > 20), "TF"]
set.seed(myseed)
rem4 = sample(extremeICs, round(length(extremeICs)*0.3))


plot_df = plot_df[!plot_df$TF %in% c(rem, rem2, rem3, rem4), ]
pwmSampled = pwmSampled[!names(pwmSampled) %in% c(rem, rem2, rem3, rem4)]
pcmSampled = pcmSampled[!names(pcmSampled) %in% c(rem, rem2, rem3, rem4)]
```

```{r, fig.width=10, fig.height=10}
grid.arrange(
  ggplot(plot_df, aes(x=len, y=IC, color=fake)) + 
    geom_point() + 
    theme_bw() +
    scale_color_manual(values=c("black","red")),
  
  ggplot(plot_df, aes(x=len, y=IC, color=fake_group)) + 
    geom_point() + 
    theme_bw() +
    scale_color_manual(values=c("black","blue",rep("orange", 7),"red")),
  
  ncol=2
)
```

```{r, fig.width=10, fig.height=10}
grid.arrange(
  ggplot(plot_df, aes(x=GC_content, y=Repeat_content, color=fake)) + 
    geom_point() + 
    theme_bw() +
    scale_color_manual(values=c("black","red")),
  
  ggplot(plot_df, aes(x=GC_content, y=Repeat_content, color=fake_group)) + 
    geom_point() + 
    theme_bw() +
    scale_color_manual(values=c("black","blue",rep("orange", 7),"red")),
  
  ncol=2
)
```

```{r, fig.width=12, fig.height=12}
#add in the fake ones when done. 
grid.arrange(
  ggplot(plot_df, aes(x=chr10_prevalence, y = len, color=fake_group)) + 
    geom_point() + theme_bw() +
    scale_color_manual(values=c("black","blue",rep("orange", 7),"red")),
  
  ggplot(plot_df, aes(x=chr10_prevalence, y = IC, color=fake_group)) + 
    geom_point() + theme_bw() +
    scale_color_manual(values=c("black","blue",rep("orange", 7),"red")),
  
  ggplot(plot_df, aes(x=chr10_prevalence, y = GC_content, color=fake_group)) + 
    geom_point() + theme_bw() +
    scale_color_manual(values=c("black","blue",rep("orange", 7),"red")),
  
  ggplot(plot_df, aes(x=chr10_prevalence, y = Repeat_content, color=fake_group)) + 
    geom_point() + theme_bw() +
    scale_color_manual(values=c("black","blue",rep("orange", 7),"red")),
  
  ncol=2
)
```


```{r}
svg("FigS5_simPWMs.svg", width = 25, height = 8)
grid.arrange(
  ggplot(plot_df, aes(x=chr10_prevalence, y = len, color=fake)) + 
    geom_point() + theme_bw() +
    theme(legend.position="none", axis.text = element_text(size = 24)) +
    xlab("") + ylab("") +
    scale_color_manual(values=c("black","orange")),
  
  ggplot(plot_df, aes(x=chr10_prevalence, y = IC, color=fake)) + 
    geom_point() + theme_bw() +
    theme(legend.position="none", axis.text = element_text(size = 24)) +
    xlab("") + ylab("") +
    scale_color_manual(values=c("black","orange")),
  
  ggplot(plot_df, aes(x=chr10_prevalence, y = GC_content, color=fake)) + 
    geom_point() + theme_bw() +
    theme(legend.position="none", axis.text = element_text(size = 24)) +
    xlab("") + ylab("") +
    scale_color_manual(values=c("black","orange")),
  ncol=3
)
dev.off()
```


```{r}
seqLogo(ppmSampled$FAKE_C10_9)
seqLogo(ppmSampled$FAKE_C6_3)
seqLogo(ppmSampled$FAKE_C10_14)
```


### 5. How similar are the fake + real ones?
```{r}
#which = "Euclidean"
which = "Pearson"

all_pwms = c(jaspar_pwm, pwmSampled)

mylist = list()
for (tf in names(all_pwms)){
  tmp = PWMSimilarity(all_pwms, all_pwms[[tf]], method=which)
  mylist[[tf]] = as.vector(tmp)
}

sim_mat = do.call(cbind, mylist)
rownames(sim_mat) = names(all_pwms)
sim_mat[1:3,1:3]
```

```{r, fig.height=20, fig.width=20}
clusters_tokeep = names(table(pwm_clusters$cluster)[table(pwm_clusters$cluster) > 30])
tfs_tokeep = rownames(pwm_clusters[pwm_clusters$cluster %in% clusters_tokeep, ])

sim_mat_sub = sim_mat[rownames(sim_mat) %in% tfs_tokeep, colnames(sim_mat) %in% tfs_tokeep]

colors <- brewer.pal(length(clusters_tokeep), "Set3")  
cluster_color <- setNames(colors, clusters_tokeep)

ha = HeatmapAnnotation(JASPAR_CLUSTER = pwm_clusters[match(colnames(sim_mat_sub), rownames(pwm_clusters)), "cluster"],
                       col = list(JASPAR_CLUSTER = cluster_color))

Heatmap(as.matrix(sim_mat_sub), 
        top_annotation = ha,
        col=colorRamp2(c(1, 0.8, 0.3, 0), c("#800080", "cornflowerblue", "white", "white")), 
        name = "Pearson Distance",
        show_row_names = FALSE,
        show_column_names = FALSE)
```

```{r, fig.height=20, fig.width=20}
sim_mat_sub = sim_mat[rownames(sim_mat) %in% names(jaspar_pwm), colnames(sim_mat) %in% names(jaspar_pwm)]
Heatmap(as.matrix(sim_mat_sub), 
        col=colorRamp2(c(1, 0.8, 0.3, 0), c("#800080", "cornflowerblue", "white", "white")), 
        name = "Pearson Distance",
        show_row_names = FALSE,
        show_column_names = FALSE)
```

```{r, fig.height=20, fig.width=20}
png("FigS5_simPWMs_heatmap.png", width = 3000, height = 1500, units = "px", res = 300)
ha = HeatmapAnnotation(simulated = ifelse(grepl("FAKE", colnames(sim_mat)), "simulated", "real"),
                       col = list(simulated = c("real" = "black", "simulated"="orange"))
)
Heatmap(as.matrix(sim_mat), 
        col=colorRamp2(c(1, 0.8, 0.3, 0), c("#800080", "cornflowerblue", "white", "white")), 
        name = "Pearson Distance",
        top_annotation = ha,
        show_row_names = FALSE,
        show_column_names = FALSE)
dev.off()
```


## Finally, output files so that footprinting can take place. 
```{r}
write_matrix_to_file <- function(i, filename) {
  matrix = pcmSampled[[i]]@profileMatrix
  
  lines <- apply(matrix, 1, function(row) {
    row_name <- rownames(matrix)[which(apply(matrix, 1, identical, y = row))]
    paste(row_name, "[", paste(sprintf("%5d", row), collapse=" "), "]")
  })
  # Use cat to append to the file
  cat(paste0(">", i, "\t", i), file = filename, sep = "\n", append = TRUE)
  cat(paste(lines, collapse = "\n"), file = filename, sep = "\n", append = TRUE)
}

# Write the output matrix to a file
for (i in names(pcmSampled)){
  write_matrix_to_file(i, "Simulated_PFMs_jaspar.txt")
}
```

```{r}
mydir = "final_simPWMs/"
dir.create(mydir, recursive = TRUE, showWarnings = FALSE)
for (i in names(pcmSampled)){
  filename = paste0(i, ".pwm")
  write.table(pcmSampled[[i]]@profileMatrix, 
              paste0(mydir, filename), 
              quote = F, row.names = F, col.names = F)
}
```

```{r}
save.image(file="checkpoint.RData")
```

```{r}
sessionInfo()
```



```{r}
#version of monaLisa::dumpJaspar

#filename="~/Desktop/JASPAR2022_CORE_vertebrates_non-redundant_pfms_homer.motif"
#siteList <- TFBSTools::readJASPARMatrix("~/projects/benchmarking/simulatedPWMs/JASPAR2022_CORE_vertebrates_non-redundant_pfms_jaspar.txt", matrixClass=c("PFM"))

load("~/checkpoint.RData")
siteList = pcmSampled
filename="~/Desktop/Simulated_PFMs_homer.motif"

pseudocount = 1
fh <- file(filename, "wb")

for (i in seq_len(length(siteList))) {
  ppm <- TFBSTools::Matrix(siteList[[i]]) + pseudocount
  ppm <- t(t(ppm)/colSums(ppm))
  tmp.rn <- rownames(ppm)
  scorecut <- log(2^10)
  ppm <- apply(ppm, 2, function(x) { sprintf("%.3f", x) })
  rownames(ppm) <- tmp.rn
  wm.name <- paste0(TFBSTools::ID(siteList[[i]]), ":::", 
                    TFBSTools::name(siteList[[i]]))
  cat(sprintf(">%s\t%s\t%.2f\t-10\n", paste(apply(ppm, 2, function(x) {
    rownames(ppm)[which.max(x)]
    }), collapse = ""), wm.name, scorecut), file = fh, append = TRUE)
  write.table(file = fh, t(ppm), row.names = FALSE, col.names = FALSE, 
              sep = "\t", quote = FALSE, append = TRUE)
  flush(fh)
}
close(fh)
    
```
